Remark on the hypothetical judgment

### 2015-08-01 paper.tex
remove citation from abstract

What is the proper explanation of intuitionistic hypothetical judgment, and
thence propositional implication? The answer to this question is not clearly
determinate from the writings of Brouwer, Heyting and Kolmogorov, since in their
lifetimes they propounded multiple (sometimes conflicting) explanations of the
hypothetical judgment \cite{van-dalen:2004}. To my mind, the determination of an
acceptable explanation must be, then, an act of abduction and reconstruction,
taking into account its adequacy for the expression of the bar theorem and, more
generally, the development of an open-ended framework for transcendental
arguments in mathematics.

@@

Now, in contrast to the treatment of the quantifiers in formal intuitionistic
logic (or in the Mitchell-B\'enabou language of a topos
\cite{maclane-moerdijk}), it is part of their meaning that their verifications
should respect the equality of the domain of discourse; this constraint is
called functionality, and reflects the fact that universal quantification is
taken as a more general form of implication. Within the theory of sets, there is
simply not a quantifier which expresses non-functional generality; in this way,
contrary to the state of affairs in \BISH{} \cite{bishop:1967}, the theorem of
choice is in fact verified by a choice \emph{function}, not merely a choice
\emph{operation}.

>>>>

What is the proper explanation of intuitionistic hypothetical judgment, and
thence propositional implication? The answer to this question is not clearly
determinate from the writings of Brouwer, Heyting and Kolmogorov, since in their
lifetimes they propounded multiple (sometimes conflicting) explanations of the
hypothetical judgment. To my mind, the determination of an
acceptable explanation must be, then, an act of abduction and reconstruction,
taking into account its adequacy for the expression of the bar theorem and, more
generally, the development of an open-ended framework for transcendental
arguments in mathematics.

@@

Now, in contrast to the treatment of the quantifiers in formal intuitionistic
logic (or in the Mitchell-B\'enabou language of a topos
\cite{maclane-moerdijk}), in this setting it is part of their meaning that their
verifications should respect the equality of the domain of discourse; this
constraint is called functionality, and reflects the fact that universal
quantification is reconstructed as a more general form of implication. Within
the theory of sets, there is simply not a quantifier which expresses
non-functional generality; in this way, contrary to the state of affairs in
\BISH{} \cite{bishop:1967}, the theorem of choice is in fact verified by a
choice \emph{function}, not merely a choice \emph{operation}.
